{DEM}-{SPH} study of particle dispersion in fluid

### 2024-08-13 manuscript/paper.tex
Update manuscript

  \title{{DEM}-{SPH} study of particles dispersion in fluid}
  \author[xxx]{Dinesh Adepu\corref{cor1}}
  \ead{d.dinesh@surrey.ac.uk}
  \author[xxx]{Chuan Yu Wu}
  % \ead{XXX}
\address[xxx]{School of Chemistry and Chemical Engineering, University of Surrey}

@@

\begin{abstract}
  Mixing powdered substances in tanks using a stirrer is a common occurrence
  across various industries. Typically, this phenomenon is studied numerically
  due to the complex nonlinear physics involved. In our current study, we have
  developed a solver combining Smoothed Particle Hydrodynamics (SPH) and
  Discrete Element Method (DEM) to investigate the mixing behavior of rigid
  circular particles within a tank under the influence of a stirrer operating
  at different speeds.  The SPH method governs the fluid phase, while the
  dynamics and interactions of particles are captured by the DEM. We achieve a
  fully resolved coupling between solid particles and fluid particles by
  discretizing solid particles into dummy SPH particles. Our validation
  process includes verifying the fluid solver's accuracy through a Poiseuille
  problem and validating the DEM solver by benchmarking interactions between
  two particles and particle-wall impacts.  The coupled model's is validated
  by simulating single-particle entry in a 2D steady tank and a cube settling
  in a tank, which is a 3D case. Following validation, we proceed with the
  mixing study. We conduct two simulations: one with spherical particles of
  uniform radius and another with particles having two different radius.  Our
  findings reveal that as for lower stirrer speeds, particles tend to
  aggregate initially and remain in the center. However, with further increase
  in stirrer speed, particles tend to accumulate near the tank corners due to
  the circulation induced in the fluid due to the stirrer motion. We use an
  open-source software, PySPH, adapting and integrating the coupled model to
  conduct the current study.
\end{abstract}

@@

Particle dispersion is a fundamental physical phenomenon encountered in
various industries where powder and fluid interactions are pivotal.  Petroleum
industry \cite{wang2023developments}, the transport of bodies in internal
systems \cite{Dai2021}, debris flow \cite{Qingyun2022}, the food processing
industry \cite{Karunasena2014}, and ice-sea modeling \cite{Mintu2018} are a
few areas to mention.  In many processes the mixing of powder with fluid using
a stirrer are commonplace \cite{li2022study}.  \Cref{fig:gloabl_body_frame_rb}
illustrates a sample particle dispersion problem studied in the current work.
While experimental investigation of such processes is often impractical due to
their highly nonlinear nature, numerical methods offer a viable alternative.
While these systems are studied with numerical methods, they are a part of
two-way coupling models. In numerical method approach, one can employ either a
mesh-based or meshless technique to study the current phenomena.

@@

The combination of Discrete Element Method (DEM) for solid particle
interactions and Computational Fluid Dynamics (CFD) for fluid modeling
is a classic approach in handling particulate flow problem. In
mesh-based modeling, lattice Boltzmann method (LBM)
\cite{xiong2014lbm} and finite volume method (FVM) \cite{kloss2012models} are
utilized to handle the fluid dynamics. Similarly, in meshless schemes like
Smoothed Particle Hydrodynamics (SPH) \cite{peng2021fully},
Moving Particle Semi-implicit (MPS), and Particle Finite Element
Method (PFEM) \cite{li2019modeling, franci2020pfem}, handles the fluid
dynamics.

While handling the interaction between the fluid and solid particles, the
interaction can be categorized into a resolved and unresolved coupling.  In
unresolved coupling, simplified models are utilized to calculate the forces
exerted on solid particles within the fluid flow. In contrast, in the fully
resolved coupling approach, the fluid forces are directly computed on the
solid particles without the need for empirical drag force models.  Both
strategies are applicable in both meshless \cite{cleary2015prediction,
  trujillo2020smooth} and mesh-based \cite{van2008numerical, ma2022review}
CFD-DEM techniques.  In addition to the study mono-sized particle behaviour in
a fluid \cite{trujillo2020smooth}, research such as
\cite{brosh2014accelerating} and \cite{peng2021fully} explore wide particle
size distributions with CFD-DEM and SPH-DEM respectively.  Regarding
non-Newtonian fluid modelling in mesh-based schemes, works like
\cite{li2018dam} are notable and in meshless schemes, \cite{peng2021fully} is
notable. \citet{zhu2008discrete} and \citet{ma2022review} delve into
investigations on spherical and non-spherical particles respectively within
CFD-DEM frameworks.  Meshless CFD-DEM, though fruitful, encounters challenges
in handling problems with free surfaces and becomes computationally demanding
with a large number of solid particles, especially in fully resolved
coupling. In contrast, SPH offers advantages in such scenarios due to its
inherent capability in handling free surfaces and adeptness in modeling large
mesh deformation problems, facilitating the incorporation of novel physics.

The SPH method is a meshless numerical method originally proposed by
\citet{gingold1977smoothed} and \citet{lucy1977numerical} to model
astrophysics problems. It has been extensively applied to simulate problems
involving fluids, structural dynamics, fluid-structure interaction, granular
physics, non-Newtonian fluid flows \cite{peng2021fully} among other areas
\cite{monaghan2012smoothed}.  Various kind of particulate flows, involving,
simple spherical particles mixing using unresolved coupling
\cite{cleary2015prediction, markauskas2019coupled} and with debris flow
\cite{canelas2016sph} with fully resolved coupling approach were analysed
using SPH-DEM.  Handling particles of arbitrary shape in fluid flow is
addressed in various works, including those by \citet{peng2021fully,
  canelas2016sph, amicarelli2015smoothed}.  Several variants of SPH, such as
WCSPH \cite{peng2021fully, cleary2015prediction}, and ISPH
\cite{asai2021fluid}, are utilized for modeling fluid flow with particulate
flows. The coupling between fluid and rigid bodies in the fully resolved
category employs techniques such as the fixed ghost particle technique
\cite{canelas2016sph, asai2021fluid}, single layers of dummy SPH particles
\cite{peng2021fully}, and simple repulsive forces
\cite{monaghan2009sph}. Coupling between SPH fluid particles and solid
structures discretized as vertex edges is also proposed by
\citet{park2023new}. Particle settling with variable sizes is studied by
\citet{zou2022study}.  Elastic behavior of spherical particles and dispersion
studies are conducted by \citet{ng2021numerical}.  Interaction between rigid
bodies of arbitrary sizes is addressed using various techniques, such as those
discussed in papers by \citet{asai2021fluid}, \citet{peng2021fully},
\citet{canelas2016sph}.

In the current work we consider the particle dispersion within a large-scale
under stirrer mixing phenomenon. We employ Smoothed Particle Hydrodynamics
(SPH) to model fluid dynamics and the Discrete Element Method (DEM) to handle
the dynamics and contact interactions among spherical particles. The
interaction between solids and fluids is modelled using a fixed ghost particle
approach, ensuring a fully resolved simulation. We validate the fluid solver
independently using Poiseuille flow and the DEM solver separately using
fundamental benchmarks such as particle-particle and particle-wall
impacts. Subsequently, the fully resolved coupled SPH-DEM solver is validated
using three benchmarks: a single particle entering, two particles settling,
and a forced wedge entry into a steady hydrostatic tank. Upon confirming the
fidelity of the solver, we proceed with analyzing the mixing behavior of
various spherical particles influenced by a stirrer in a tank. We investigate
scenarios involving particles of different densities, stirrer velocities, and
variable sizes as part of our case study. The development work is conducted
using PySPH\cite{ramachandran2021pysph}, an open-source code available at
\url{https://github.com/pypr/pysph}.  We modify the current codebase to
integrate implementations of the Discrete Element Method (DEM) and the
Smoothed Particle Hydrodynamics (SPH) coupled with DEM model for our ongoing
research. To ensure reproducibility, we utilize the automan package
\cite{ramachandran2018automan} to automate all results generated in the
current manuscript, aiming for a reproducible research approach.

The structure of this paper is as follows: In
\cref{sec:fluid-modeling}, we explain the numerical method used to model fluid
dynamics. \Cref{sec:rbd} outlines the equations governing the dynamics
of rigid bodies and the contact force model employed to resolve collisions
among them. The coupling between rigid spherical particles and the fluid is
detailed in \cref{sec:rfc}. Our results are presented in
\cref{sec:results}, where various problems from the literature are simulated to
validate the current solver, and particle dispersion is examined.

\FloatBarrier%
\section{Fluid modeling}
\label{sec:fluid-modeling}
In the current work, we follow a weakly-compressible SPH approach to model the
fluid. The continuum in SPH is modeled using particles, which has physical
properties such as mass, velocity, and these particles interact based on the
governing equations using a Guassian-like kernel
\cite{monaghan-review:2005,morris1997modeling}.

@@

The stress tensor is split into pressure and viscous parts,
\begin{equation}
  \label{eq:fluid-stress-decomposition}
  \sigma_{ij} = - p \delta_{ij} + 2 \eta \frac{\partial u_i}{\partial x_j}
\end{equation}
where $\eta$ is the kinematic viscosity of the fluid, $p$ is the pressure,
$\delta_{ij}$ is the Kronecker delta function.

\FloatBarrier%
\subsection{Discretized fluid governing equations}
\label{sec:sph--governing-equations}
The governing equations involve function, derivative and divergence
operators. In SPH, these operators are approximated based on the positions,
mass and the kernel values of the discretized particles. Assume that the
domain is discretized into N particles with mass $m_i$ and volume
$V_i$, We have
\begin{equation}
  \label{eq:mass_repr}
  m_i = V_i \> \rho_i.
\end{equation}
Based on this discretization, the discrete function approximation is given as,
\begin{equation}
  \label{eq:discrete_form}
  A_i = \sum_{j \in \text{Neigh}(i)}\> \frac{m_j}{\rho_j} A_j\> W(\ten{x}_i - \ten{x}_j, h),
\end{equation}
where $A(\boldsymbol{x}_i) = A_i$ is the value of the field property of
particle i, similarly for $A_j$. Here, $\text{Neigh}(i)$ is the neighbours of
particle $i$.  A symmetric derivative approximation~\cite{Violeau16} is given
as,
\begin{equation}
  \nabla A(\ten{x}_i) = \rho_i \sum_{j} m_j \left(\frac{A_i}{\rho_i^2} + \frac{A_j}{\rho_j^2}\right) \nabla W_{ij}.
\end{equation}
A symmetric gradient operator ensures equal and opposite forces acting on the
particles.

An approximation of the divergence operator is used to discretize the
\cref{eq:ce}.  It is given as,
\begin{equation}
  \label{intro:eq:sph-continu-div-final}
  \nabla \cdot \ten{v}(\ten{x}_i) = \frac{1}{\rho_i} \sum_{j} m_j \left(\ten{v}_j - \ten{v}_i\right) \cdot \nabla W_{ij}.
\end{equation}

% \subsection{Discretized fluid governing equations}
The SPH discretization of the continuity
equation~\cref{eq:sph-discretization-continuity} is given as,
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \;
  \rho_{a} \; {\ten{u}}_{ab} \; \cdot \nabla_{a} W_{ab},
\end{equation}

%
Similarly, the discretized momentum equation is written as,
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg)
  \nabla_{a} W_{ab}
 \;+\;
  \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab}  \;+\;
  \ten{g}_{a},
\end{multline}
where $\ten{I}$ is the identity matrix, $\eta$ is the kinematic viscosity of the
fluid. The viscous term in \Cref{eq:me} is discretized according to the
formulation introduced in \cite{morris1997modeling}.

We add to the momentum equation an additional artificial viscosity term
$\Pi_{ab}$~\cite{monaghan-review:2005} to maintain the stability of the
numerical scheme, given as,
\begin{align}
  \label{eq:mom-av}
  \Pi_{ab} =
  \begin{cases}
\frac{-\alpha h_{ab} \bar{c}_{ab} \phi_{ab}}{\bar{\rho}_{ab}}
  & \ten{u}_{ab}\cdot \ten{r}_{ab} < 0, \\
  0 & \ten{u}_{ab}\cdot \ten{r}_{ab} \ge 0,
\end{cases}
\end{align}
where,
%
\begin{equation}
  \label{eq:av-phiij}
  \phi_{ab} = \frac{\ten{u}_{ab} \cdot \ten{r}_{ab}}{r^2_{ab} + 0.01 h^2_{ab}},
\end{equation}
%
where $\ten{r}_{ab} = \ten{r}_a - \ten{r}_b$,
$\ten{u}_{ab} = \ten{u}_a - \ten{u}_b$, $h_{ab} = (h_a + h_b)/2$,
$\bar{\rho}_{ab} = (\rho_a + \rho_b)/2$, $\bar{c}_{ab} = (c_a + c_b) / 2$, and
$\alpha$ is the artificial viscosity parameter.  The pressure $p_a$ is evaluated
using an equation of state:
\begin{equation}
\label{eqn:sph-eos}
  p_a = K \bigg(\frac{\rho_a}{\rho_{0}} - 1 \bigg).
\end{equation}
Where, $K=\rho_0 \, c_0^2$ is bulk modulus of the body, with
$c_0=10 \times V_{\text{max}}$ is speed of sound, while $\rho_0$ as the
initial density of the particles.

@@

The ghost particle approach proposed by \citet{Adami2012} is used to model the
boundaries. We use three layers of ghost particles to model the solid wall.
The properties of the solid wall are interpolated from the fluid particles.

When the viscous force is computed, the no slip boundary condition is used,
where the velocity on the boundary set as,
\begin{equation}
  \label{eq:no-slip-bc-u}
  \ten{u}_{\text{Ga}} = 2 \ten{u}_{\text{p}} - \ten{\hat{u}}_{\text{a}}.
\end{equation}
The projected velocity $\ten{\hat{u}}_{\text{a}}$ on the ghost particles is
computed using,
\begin{equation}
  \label{eq:v-ghost}
  \ten{\hat{u}}_a = \frac{\sum_b\ten{u}_b W_{ab}}{\sum_b W_{ab}},
\end{equation}
where $\ten{u}_b$, is the velocity of the fluid particle $b$ and $W_{ab}$ is the kernel
value between the fluid particle and the ghost particle.

@@

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.7\textwidth]{images/spherical_particles_dem_representation}
  \caption{Demonstration of contact handling between two rigid spherical
    particles immersed in a fluid tank.}
  \label{fig:spherical-particles-in-tank-dem}
\end{figure}

>>>>

  \title{{DEM}-{SPH} study of particle dispersion in fluid}
  \author[xxx]{Dinesh Adepu\corref{cor1}}
  \ead{d.adepu@surrey.ac.uk}
  \author[xxx]{Chuan Yu Wu}
  % \ead{XXX}
\address[xxx]{School of Chemistry and Chemical Engineering, University of Surrey, Guildford, GU2 7XH, UK}

@@

\begin{abstract}
  Mixing powdered substances in stirring tanks is a common occurrence across
  various industries. In our current study, we developed a solver combining
  Smoothed Particle Hydrodynamics (SPH) and Discrete Element Method (DEM) to
  investigate the mixing behavior of particles in a tank under the influence
  of a stirrer operating at different speeds.  The SPH method governs the
  fluid phase, while the dynamics and interactions of particles are captured
  by DEM. We achieve a fully resolved coupling between solid particles and
  fluid particles by discretizing solid particles into dummy SPH
  particles. Validation was performed, including verifying the fluid solver's
  accuracy through a Poiseuille problem and validating the DEM solver by
  benchmarking interactions between two particles and particle-wall impacts.
  The coupled model is validated by simulating single-particle entry in a
  steady tank in 2D and a cube settling in a tank in 3D. Using the validated
  model, we conduct two simulations: one with spherical particles of uniform
  radius and another with particles having two different radius.  Our results
  reveal that at lower stirrer speeds, particles tend to aggregate
  initially and remain in the center. However, with the increase in
  stirrer speed, particles tend to accumulate near the tank corners due to the
  circulation induced in the fluid due to the stirrer motion.
\end{abstract}

@@

Particle dispersion is a fundamental physical phenomenon encountered in
various industries where powder and fluid interactions are pivotal.  Control
of water flooding in petroleum industry \cite{wang2023developments}, the
transport of bodies in internal systems \cite{Dai2021}, debris flow
\cite{Qingyun2022}, the food processing industry \cite{Karunasena2014}, and
ice-sea modeling \cite{Mintu2018} are a few examples to mention.  In many
processes, the mixing of powder with fluid using a stirrer are commonplace
\cite{li2022study}.  \Cref{intro:schematic} illustrates a sample particle
dispersion problem.  While experimental investigation of such processes is
often impractical due to their highly nonlinear nature, numerical methods
offer a viable alternative.  While these systems are studied with numerical
methods, they are generally treated as a two-way coupling problem. In
numerical method approach, one can employ either a mesh-based or meshless
technique to study these phenomena.

@@

The combination of Discrete Element Method (DEM) for solid particle
interactions and Computational Fluid Dynamics (CFD) for fluid modeling is a
classic approach in handling particulate flow problem. In mesh-based modeling,
lattice Boltzmann method (LBM) \cite{xiong2014lbm} and finite volume method
(FVM) \cite{kloss2012models} are utilized to handle the fluid
dynamics. Similarly, in meshless schemes, Smoothed Particle Hydrodynamics
(SPH) \cite{peng2021fully}, Moving Particle Semi-implicit (MPS), and Particle
Finite Element Method (PFEM) \cite{li2019modeling, franci2020pfem}, are
adopted to handle the fluid dynamics.

While handling the interaction between the fluid and solid particles, the
interaction can be categorized into a resolved and unresolved coupling.  In
unresolved coupling, simplified models are utilized to calculate the forces
exerted on solid particles within the fluid flow. In contrast, in the fully
resolved coupling approach, the fluid forces are directly computed on the
solid particles without the need for empirical drag force models.  Both
strategies are applicable in both meshless \cite{cleary2015prediction,
  trujillo2020smooth} and mesh-based \cite{van2008numerical, ma2022review}
techniques.  In addition to study mono-sized particle behaviour in a fluid
\cite{trujillo2020smooth}, research, such as \cite{brosh2014accelerating} and
\cite{peng2021fully}, also explore wide particle size distributions with
CFD-DEM and SPH-DEM respectively.  Regarding non-Newtonian fluid modelling in
mesh-based schemes, works like \cite{li2018dam} are notable and in meshless
schemes, \cite{peng2021fully} is notable. \citet{zhu2008discrete} and
\citet{ma2022review} delve into investigations on spherical and non-spherical
particles respectively within CFD-DEM frameworks.  Mesh based CFD-DEM, though
fruitful, encounters challenges in handling problems with free surfaces and
becomes computationally demanding with a large number of solid particles,
especially in fully resolved coupling. In contrast, SPH offers advantages in
such scenarios due to its inherent capability in handling free surfaces and
adeptness in modeling large mesh deformation problems, facilitating the
incorporation of novel physics.

The SPH method is a meshless numerical method originally proposed by
\citet{gingold1977smoothed} and \citet{lucy1977numerical} to model
astrophysics problems. It has been extensively applied to simulate problems
involving fluids, structural dynamics, fluid-structure interaction, granular
physics, non-Newtonian fluid flows \cite{peng2021fully} among other areas
\cite{monaghan2012smoothed}.  Various kind of particulate flows, including,
simple spherical particles mixing using unresolved coupling
\cite{cleary2015prediction, markauskas2019coupled} and with debris flow
\cite{canelas2016sph} with fully resolved coupling approach, were analysed
using SPH-DEM.  Handling particles of arbitrary shape in fluid flow is
addressed in various works, including those by \citet{peng2021fully,
  canelas2016sph}, and \citet{amicarelli2015smoothed}.  Several variants of
SPH, such as weakly compressible SPH (WCSPH) \cite{peng2021fully,
  cleary2015prediction}, and ISPH \cite{asai2021fluid}, are utilized for
modeling fluid flow with particulate flows. The coupling between fluid and
rigid bodies in the fully resolved category employs techniques such as the
fixed ghost particle technique \cite{canelas2016sph, asai2021fluid}, single
layers of dummy SPH particles \cite{peng2021fully}, and simple repulsive
forces \cite{monaghan2009sph}. Coupling between SPH fluid particles and solid
structures discretized as vertex edges is also proposed by
\citet{park2023new}. Particle settling with variable sizes is studied by
\citet{zou2022study}.  Elastic behavior of spherical particles and dispersion
studies are conducted by \citet{ng2021numerical}. Interaction between rigid
bodies of arbitrary sizes is addressed using various techniques. For example,
\citet{asai2021fluid} utilize an energy-tracking impulse method, while
\citet{peng2021fully} apply a surface mesh-based Discrete Element Method
(SMR-DEM) to handle collisions.

Mixing of particles in a tank with stirrer is modelled with an unresolved
SPH-DEM model by \citet{zhu2021virtual}. While the interaction between two
solid particles is handled using a linear model. A similar study is carried
out to study the mixing of particles in a rotating drum by \citet{he2018gpu}
using unresolved SPH-DEM model. Here, the authors use non-linear contact force
model, which captures the force model accurately.  While using a fully
resolved SPH-DEM coupling solver, particle dispersion has been studied in
\cite{ng2021numerical,canelas2016sph,canelas2017resolved,peng2021fully}, where
the interaction between the solid particle and the fluid is handled using
ghost particles placed on the solid particle. In works of unresolved coupling,
the interaction among the solid particles is handled using the discretized
ghost points with a modified DEM model. However, fully resolved coupled works
have only been applied to works with arbitrary shaped bodies, and to mainly
ocean engineering problems \cite{canelas2016sph,canelas2017resolved} but not
for particle mixing in tank.  In the current work, we study the circular
particle mixing in stirring tanks.  We employ a fully resolved coupled SPH-DEM
model, however, unlike previous fully resolved approaches
\cite{ng2021numerical,canelas2016sph,canelas2017resolved,peng2021fully}, we
only utilize the discretized ghost particles placed on the circular solid
particle for interaction with the fluid, but not for solid solid
interaction. The solid-solid interaction is carried out using the traditional
DEM methodology \cite{cundall_discrete_1979}. We validate the fluid solver
independently using Poiseuille flow and the DEM solver separately using
fundamental benchmarks such as particle-particle and particle-wall
impacts. Subsequently, the fully resolved coupled SPH-DEM solver is validated.

Upon confirming the fidelity of the solver, we proceed with analyzing the
mixing behavior of various spherical particles influenced by a stirrer in a
tank. We investigate scenarios involving particles of different densities,
stirrer velocities, and variable sizes as part of our case study. The
development work is conducted using PySPH\cite{ramachandran2021pysph}, an
open-source code available at \url{https://github.com/pypr/pysph}.  We modify
the current codebase to integrate implementations of the DEM and the SPH
coupled with DEM model for our ongoing research. To ensure reproducibility, we
utilize the automan package \cite{ramachandran2018automan} to automate all
results generated in the current manuscript, aiming for a reproducible
research approach.

% \begin{itemize}
% \item
% \item \cite{zhu2021virtual} uses unresolved coupling, with contact between the
%   particles being a linear model. Mixing in a tank with stirrer.
% \item \cite{he2018gpu} uses unresolved coupling, with contact between the
%   particles being a Hertz model, considering the real material properties of
%   the interacting particles. Rotating drum mixing.
% \item tong2018scale uses unresolved coupling, with contact between the
%   particles being a linear model. Studies mixing in tank.
% \item Circular particle dispersion study SPH-DEM-VCPM studies
%   \citet{ng2021numerical}, Fully resolved coupling, with particles interacting
%   among themselves using a linear force model, the interaction between two
%   solid particles is handled using the dummy points of VCPM.
% \item Circular particle dispersion study SPH-DEM-VCPM studies
%   \citet{ng2021numerical}, Fully resolved coupling, with particles interacting
%   among themselves using a linear force model, the interaction between two
%   solid particles is handled using the dummy points of VCPM.
% \item canelas2016sph, canelas2017resolved, peng2021fully model transport of
%   rigid body of circular and arbitrary shape in a fluid flow, mainly applied to
%   areas of ocean engineering and not anything related to spherical particle
%   mixing.  \citet{ng2021numerical} assumes the particles to be elastic and
%   models it with VCPM method, and those discretized VCPM notes also acts as
%   SPH dummy particles and rigid fluid coupling is carried out using that.
% \item Though fully resolved coupling works have been existing in the
%   literature the interaction among the solid particles are handled through the
%   dummy particles being discretized. Using the ghost particles, the contact
%   force between the solid particles has been handled, where both linear and
%   nonlinear contact force models are studied. While the contact force
%   coefficients are calibrated for the distributed contact (DCDEM).
% \item In the current work, we develop a coupled SPH-DEM solver, to model the
%   particle dispersion within a large-scale under stirrer mixing phenomenon.
% \item We adopt a fully coupled approach to model the interaction between the
%   solid particles and the fluid unlike the unresolved studies.
% \item The interaction between the solid particles is handled using the DEM
%   model, however, we will use the center of the solid bodies to handle the
%   contact, and not the dummy particles as used in \cite{}canelas2016sph,
%   \cite{canelas2017resolved}, \cite{peng2021fully}. This provides us with more accurate
%   contact force modeling, with a nonlinear force model.
% \end{itemize}

% In the current work we consider the particle dispersion within a large-scale
% under stirrer mixing phenomenon. We employ Smoothed Particle Hydrodynamics
% (SPH) to model fluid dynamics and the Discrete Element Method (DEM) to handle
% the dynamics and contact interactions among spherical particles. The
% interaction between solids and fluids is modelled using a fixed ghost particle
% approach, ensuring a fully resolved simulation. We validate the fluid solver
% independently using Poiseuille flow and the DEM solver separately using
% fundamental benchmarks such as particle-particle and particle-wall
% impacts. Subsequently, the fully resolved coupled SPH-DEM solver is validated
% using three benchmarks: a single particle entering, two particles settling,
% and a forced wedge entry into a steady hydrostatic tank. Upon confirming the
% fidelity of the solver, we proceed with analyzing the mixing behavior of
% various spherical particles influenced by a stirrer in a tank. We investigate
% scenarios involving particles of different densities, stirrer velocities, and
% variable sizes as part of our case study. The development work is conducted
% using PySPH\cite{ramachandran2021pysph}, an open-source code available at
% \url{https://github.com/pypr/pysph}.  We modify the current codebase to
% integrate implementations of the Discrete Element Method (DEM) and the
% Smoothed Particle Hydrodynamics (SPH) coupled with DEM model for our ongoing
% research. To ensure reproducibility, we utilize the automan package
% \cite{ramachandran2018automan} to automate all results generated in the
% current manuscript, aiming for a reproducible research approach.

The structure of this paper is as follows: In \cref{sec:fluid-modeling}, we
explain the numerical method used to model fluid dynamics. \Cref{sec:rbd}
outlines the equations governing the dynamics of rigid bodies and the contact
force model employed to resolve collisions among them. The coupling between
rigid spherical particles and the fluid is detailed in
\cref{sec:rfc}. \cref{sec:integration} discusses the time integration scheme
employed in the current work.  \cref{sec:validation}, where various problems
from the literature are simulated to validate the current solver.  Mixing
behaviour of spherical particles in a stirring tank is examined in
\Cref{sec:results}.

\FloatBarrier%
\section{Fluid modeling}
\label{sec:fluid-modeling}
In the current work, we follow a weakly-compressible SPH approach to model the
fluid. The continuum in SPH is modeled using particles, which have physical
properties such as mass, velocity, and these particles interact based on the
governing equations using a Guassian-like kernel
\cite{monaghan-review:2005,morris1997modeling}.

@@

The stress tensor is split into pressure and viscous parts,
\begin{equation}
  \label{eq:fluid-stress-decomposition}
  \sigma_{ij} = - p \delta_{ij} + 2 \eta \frac{\partial u_i}{\partial x_j}
\end{equation}
where $\eta$ is the dynamic viscosity of the fluid, $p$ is the pressure,
$\delta_{ij}$ is the Kronecker delta function.

\FloatBarrier%
\subsection{Discretized fluid governing equations}
\label{sec:sph--governing-equations}
The governing equations involve function, derivative and divergence
operators. In SPH, these operators are approximated based on the positions,
mass and the kernel values of the discretized particles. Assume that the
domain is discretized into N particles with mass $m_a$ and volume
$V_a$, We have
\begin{equation}
  \label{eq:mass_repr}
  m_a = V_a \> \rho_a.
\end{equation}
Based on this discretization, the discrete function approximation is given as,
\begin{equation}
  \label{eq:discrete_form}
  A_a = \sum_{b \in \text{Neigh}(a)}\> \frac{m_b}{\rho_b} A_b\> W(\ten{x}_a - \ten{x}_b, h),
\end{equation}
where $A(\boldsymbol{x}_a) = A_a$ is the value of the field property of
particle $a$, similarly for $A_b$. $\text{Neigh}(a)$ is the neighbours of
particle $a$ and $W(\ten{x}_a - \ten{x}_b, h)$ is the kernel value, with $h$
being the smoothing length. In the current work we used Quintic
spline~\cite{Violeau16} in all our simulation cases.  A symmetric derivative
approximation~\cite{Violeau16} is given as,
\begin{equation}
  \nabla A(\ten{x}_a) = \rho_a \sum_{b \in \text{Neigh}(a)} m_b \left(\frac{A_a}{\rho_a^2} + \frac{A_b}{\rho_b^2}\right) \nabla W_{ab}.
\end{equation}
A symmetric gradient operator ensures equal and opposite forces acting on the
particles.

% \subsection{Discretized fluid governing equations}
The SPH discretization of the continuity
equation~\cref{eq:ce} is given as,
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \;
  \rho_{a} \; {\ten{u}}_{ab} \; \cdot \nabla_{a} W_{ab},
\end{equation}
where $\ten{u}_{ab} = \ten{u}_a - \ten{u}_b$.

%
Similarly, the discretized momentum equation is written as,
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2} + \Pi_{ab}\bigg) \ten{I} \cdot
  \nabla_{a} W_{ab}
 \;+\;
  \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab}  \;+\;
  \ten{g}_{a},
\end{multline}
where $\ten{I}$ is the identity matrix. The viscous term in \Cref{eq:me} is discretized according to the
formulation introduced in
\cite{morris1997modeling}. $\Pi_{ab}$~\cite{monaghan-review:2005} is an
artificial viscosity item to maintain the stability of the numerical scheme,
given as,
\begin{align}
  \label{eq:mom-av}
  \Pi_{ab} =
  \begin{cases}
\frac{-\alpha h_{ab} \bar{c}_{ab} \phi_{ab}}{\bar{\rho}_{ab}}
  & \ten{u}_{ab}\cdot \ten{r}_{ab} < 0, \\
  0 & \ten{u}_{ab}\cdot \ten{r}_{ab} \ge 0,
\end{cases}
\end{align}
where,
%
\begin{equation}
  \label{eq:av-phiij}
  \phi_{ab} = \frac{\ten{u}_{ab} \cdot \ten{r}_{ab}}{r^2_{ab} + 0.01 h^2_{ab}},
\end{equation}
%
where $\ten{r}_{ab} = \ten{r}_a - \ten{r}_b$,
$\ten{u}_{ab} = \ten{u}_a - \ten{u}_b$, $h_{ab} = (h_a + h_b)/2$,
$\bar{\rho}_{ab} = (\rho_a + \rho_b)/2$, $\bar{c}_{ab} = (c_a + c_b) / 2$, and
$\alpha$ is the artificial viscosity parameter, is taken to be $0.1$ in the current
work. The pressure $p_a$ is evaluated using an equation of state:
\begin{equation}
\label{eqn:sph-eos}
  p_a = \rho_0 \, c_0^2 \bigg(\frac{\rho_a}{\rho_{0}} - 1 \bigg).
\end{equation}
where, $c_0=10 \times V_{\text{max}}$ is speed of sound, while $\rho_0$ as the
initial density of the particles.

@@

The ghost particle approach proposed by \citet{Adami2012} is used to model the
boundaries. To ensure kernel completion of the fluid particle near the
boundary, we use three layers of ghost particles to model the solid wall.  The
properties of the solid wall are interpolated from the fluid particles.

When the viscous force is computed, the no slip boundary condition is used,
where the velocity ($\ten{u}_{\text{Ga}}$) on the boundary set as,
\begin{equation}
  \label{eq:no-slip-bc-u}
  \ten{u}_{\text{Ga}} = 2 \ten{u}_{\text{p}} - \ten{\hat{u}}_{\text{a}}.
\end{equation}
The projected velocity $\ten{\hat{u}}_{\text{a}}$ on the ghost particles is
computed using,
\begin{equation}
  \label{eq:v-ghost}
  \ten{\hat{u}}_a = \frac{\sum_b\ten{u}_b W_{ab}}{\sum_b W_{ab}},
\end{equation}
where $\ten{u}_b$, is the velocity of the fluid particle $b$ and $W_{ab}$ is the kernel
value between the fluid particle and the ghost particle.

@@

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.7\textwidth]{images/spherical_particles_dem_representation}
  \caption{Illustration of contact modelling between two rigid spherical
    particles immersed in a fluid tank.}
  \label{fig:spherical-particles-in-tank-dem}
\end{figure}
